\documentclass[11pt,a4paper]{article}
\usepackage{aas_macros}

\usepackage{natbib}
% \renewcommand{\bibnumfmt}[1]{{#1}.}

\usepackage{bibunits}
\defaultbibliographystyle{unsrtnat}

\usepackage[LabelsAligned,openbib,ManyBibs]{currvita}
% \renewcommand*{\cvheadingfont}{\large\bfseries}
\renewcommand*{\cvlistheadingfont}{\Large\scshape}

\usepackage[left=1in,right=1in, bottom=1.2in]{geometry}
\usepackage{amssymb, multicol}

\usepackage{enumitem}

\usepackage[usenames,dvipsnames]{xcolor} % Custom colors
\usepackage[bookmarks, colorlinks, breaklinks, pdftitle={Cail Daley - vita},pdfauthor={Cail Daley}]{hyperref}
\hypersetup{linkcolor=blue,citecolor=blue,filecolor=black,urlcolor=MidnightBlue}

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{Cail Daley}
\rhead{Curriculum Vitae}
\cfoot{\thepage}
\setlength{\headheight}{13.6pt}

%~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~

\begin{document}
\thispagestyle{empty}

\begin{cv}{Cail M. Daley}
  \noindent \rule{0.5\linewidth}{0.4pt}

  \setlength{\cvlabelwidth}{0mm}
  \setlength{\cvlabelsep}{0mm}
  \begin{cvlist}{}
  \item ~\vspace{-2\baselineskip}%%
    \begin{multicols}{2}
    \item 
      Départment d’Astrophysique \\
      CEA Paris-Saclay \\
      Orme des Merisiers, Building 709 \\
      F-91191 Gif-sur-Yvette, France

      \columnbreak

    \item \ \\
      \textbf{Email:} \ \ \ \href{mailto:cail.daley@cea.fr}{cail.daley@cea.fr}\\
      \textbf{Phone:} \ \ \texttt{+33 07 75 72 18 07}\\ % Your phone number
      \textbf{GitHub:} \href{https://github.com/cailmdaley} {https://github.com/cailmdaley}
    \end{multicols}
  \end{cvlist}

  ~\vspace{-2\baselineskip}%%

  \noindent \rule{\linewidth}{0.4pt}

  \begin{cvlist}{Research Interests}
  \item ~\vspace{-2\baselineskip}%%
    \begin{multicols}{2}

      \begin{itemize}[label={$\vcenter{\hbox{\scriptsize$\bullet$ \ }}$}]
        \item Agentic science methodology
        \item CMB lensing
        \item Cosmic shear
        \item Large-scale structure
      \end{itemize}

      \columnbreak

      \begin{itemize}[label={$\vcenter{\hbox{\scriptsize$\bullet$ \ }}$}]
        \item Kinematic lensing
        \item CMB $\times$ galaxy cross-correlations
        \item Verification of AI-mediated research
        \item Weak lensing systematics
      \end{itemize}
    \end{multicols}
  \end{cvlist}

  \setlength{\cvlabelsep}{1em}
  \settowidth{\cvlabelwidth}{\cvlabelfont 2024--Present}%
  \begin{cvlist}{{Current Position}}
  \item [2024--Present]
    \textbf{Postdoctoral Researcher} \\
    CEA Paris-Saclay, CosmoStat Laboratory \\
    $\cdot$ Conducting cosmological research using agentic AI workflows, with analyses in weak gravitational lensing (UNIONS, Euclid) and CMB lensing (SPT). \\
    $\cdot$ Leading SPT$\times$Euclid Memorandum of Understanding for CMB lensing cross-correlations.
  \end{cvlist}

  % \settowidth{\cvlabelwidth}{\cvlabelfont 2018--2024}%
  \begin{cvlist}{{Education}}
  \item[2018--2024]
    \textbf{Ph.D.  \textbar{}} Astronomy \\
    University of Illinois Urbana-Champaign (UIUC) \\
    Thesis: \textit{CMB Lensing Measurements with Two Years of Data from the SPT-3G Survey}

  \item[2014--2018]
    \textbf{B.A.} \textbar{} Astronomy \& Physics (Dual Major) \\
    {Wesleyan University}, Middletown, CT \\
    Honors Thesis: \textit{Using Vertical Structure to Infer the Dynamical Mass Hidden in the AU Mic Debris Disk}
  \end{cvlist}

  % \settowidth{\cvlabelwidth}{\cvlabelfont 08/2018--07/2024}%
  \begin{cvlist}{Research Experience}
  \item[2024--Present]
    \textbf{CEA Paris-Saclay} \textbar{} CosmoStat Laboratory \\
    $\cdot$ Led end-to-end cosmic shear analysis of the UNIONS survey using agentic AI workflows ($\sim$58,000 lines of agent-written Python code). \\
    $\cdot$ Principal postdoc in Euclid CMB cross-correlations working group; leading SPT$\times$Euclid MOU. \\
    $\cdot$ Supervised three master's students on kinematic lensing projects conducted with agentic methods. \\
    $\cdot$ Developed verification practices for AI-mediated scientific analysis.

  \item[2019--2024]
    \textbf{University of Illinois Urbana-Champaign} \textbar{} advised by Gil Holder \\
    $\cdot$ Member of the South Pole Telescope project, studied gravitational lensing of the CMB and mm-wavelength transients.
    Lead one of three pipelines in the SPT-3G 2019+2020 CMB lensing analysis, producing the deepest CMB lensing maps to date.

  \item[2014--2018]
    \textbf{Wesleyan University} \textbar{} advised by A. Meredith Hughes \\
    $\cdot$ Studied debris disks at sub-mm wavelengths, led reduction and analysis of observations of the debris disk around AU Mic to measure its vertical structure at mm wavelengths for the first time.

  \item[2016--2017]
    \textbf{Leiden University} \textbar{} advised by Catherine Walsh \\
    $\cdot$ Modeled the kinematic structure of the circumstellar disk HD 100546. Selected for Leiden's LEAPS program from a pool of 450 applicants.
  \end{cvlist}

  \setlength{\cvlabelsep}{2em}
  \begin{cvlist}{International Collaborations}
  \item[2024--Present] \textbf{Euclid Consortium} --- Principal postdoc in CMB cross-correlations working group. Leading SPT$\times$Euclid Memorandum of Understanding.
  \item[2024--Present] \textbf{UNIONS Collaboration} --- Leading cosmic shear analysis for first data release.
  \item[2018--Present] \textbf{South Pole Telescope (SPT-3G)} --- Led one of three CMB lensing reconstruction pipelines. Helped colleagues adopt agentic workflows for CMB analyses.
  \end{cvlist}

  \begin{cvlist}{Outreach}
  \item[2020-2023] \textbf{SPT First Discoveries:} Participated in SPT's premier outreach program that brings together pre-K and elementary students with scientists in classrooms to encourage participation in science from a young age. Planned and remotely led $~\sim$5 lessons at the predominantly-Black Fiske Elementary in Chicago.

  \item[2020-2022] \textbf{UIUC Astronomy on Tap:} Organized monthly all-ages outreach events featuring conversations between astronomers and the public in informal settings. Involved in all aspects of program--giving talks, finding speakers, venue booking, etc. Livestreamed on \href{https://www.youtube.com/@astroontapcu219/streams}{YouTube} during the pandemic, and held in-person at several establishments in the Urbana-Champaign area.
  \item[2019] \textbf{UIUC Education Justice Project:} Worked with the university's college-in-prison program and led a workshop on programming and astronomy data analysis at the Danville Correctional Center.
  \end{cvlist}

  \begin{cvlist}{Supervision \& Mentorship}
  \item[2024--2025] \textbf{Master's Student Supervision} (CosmoStat, CEA Paris-Saclay). Supervised three master's students on kinematic lensing feasibility studies using spectroscopic survey data (GAMA, DESI). Students trained in agentic analysis workflows: designing analyses, specifying verification checks, and interpreting agent-produced results.

  \item[2022--2023] \textbf{Research Mentor:} Zimo Qu (undergraduate, transferred to UC Berkeley). Supported student in a search for stellar flares observed simultaneously by the SPT and TESS telescopes. Zimo presented a poster at the 2023 Illinois Astrofest.

  \item[2020--2021] \textbf{Undergraduate Mentor:} Society for Equity in Astronomy mentorship program. Monthly meetings to discuss research, graduate school, and other topics. Mentored three students.
  \end{cvlist}

  \begin{cvlist}{Teaching \& Service}
  \item[2021-2022] \textbf{UIUC Astronomy Journal Club:} Organized a weekly journal club with graduate students giving talks on their work and recent papers in the field.

  \item[2019] \textbf{Teaching Assistant,} ASTR 122: Stars and Galaxies

  \item[2018] \textbf{Teaching Assistant,} ASTR 404: Stellar Astrophysics
  \end{cvlist}

  \begin{cvlist}{Schools \& Workshops}
  \item[2020] Michigan Cosmology School \textit{(virtual)}

  \item[2019] {La Serena School for Data Science}, La Serena, Chile

  \item[2019] {Open Science Grid School}, Madison, WI
  \end{cvlist}

  \newpage

  \settowidth{\cvlabelwidth}{\cvlabelfont 2018-2019}%
  \setlength{\cvlabelsep}{3em}
  \begin{cvlist}{Awards}

  \item[2023] \textbf{APS DAP Student/Early Career Meeting Award} (\$600) \\
    $\cdot$ Travel to APS April Meeting 2023, Minneapolis, MN

  \item[2021] \textbf{Chambliss Astronomy Achievement Student Award} \\
    $\cdot$ AAS 238th Meeting (\textit{virtual})

  \item[2019-2020] \textbf{Center for Astrophysical Surveys Graduate Fellowship}  (\$30,000) \\
    $\cdot$ University of Illinois Urbana-Champaign

  \item[2018] \textbf{Thesis High Honors} \\
    $\cdot$ Wesleyan University

  \item[2017] \textbf{Student Travel Grant} (\$1000) \\
    $\cdot$ NASA Connecticut Space Grant Consortium\\
    $\cdot$ Travel to AAS

  \item[2017]
    \textbf{3rd Prize, Visualizing Knowledge Exhibition} \\
    $\cdot$ Wesleyan University\\
    $\cdot$ Title: \textit{Orbital Motion of Gas in Planetary System HD 100546}

  \item[2017]
    \textbf{Siver Scholarship}\\
    $\cdot$ Wesleyan University\\
    $\cdot$ awarded to undergraduate students majoring in or demonstrating strong academic interest in physics

  \item[2015] \textbf{Undergraduate Research Fellowship} (\$5000) \\
    $\cdot$ NASA Connecticut Space Grant Consortium \\
    $\cdot$ Title: \textit{Searching for Non-Axisymmetry in the Unusual Gas Disk Around a Main Sequence Star}

  \item[2015] \textbf{Research in Sciences Fellowship} (\$4000)\\
    $\cdot$ Wesleyan University\\
    $\cdot$ Title: \textit{Searching for Non-Axisymmetry in the Unusual Gas Disk Around a Main Sequence Star}

  \end{cvlist}

  \begin{cvlist}{Skills}
  \item \
  \item \textit{Languages \& tools:} Python (advanced), Unix/Linux (advanced), git, \LaTeX, Snakemake, Julia (intermediate).
  \item \textit{Cosmology software:} CAMB, HEALPix, NaMaster, lenspyx, TreeCorr, emcee, CosmoSIS.
  \item \textit{AI \& computing:} Agentic scientific workflows (Claude, frontier LLMs), Bayesian inference, machine learning, high-performance computing (Jean Zay, SLURM), workflow orchestration.

  \end{cvlist}

  \newpage

  \settowidth{\cvlabelwidth}{\cvlabelfont 01/2024}%
  \begin{cvlist}{Talks \& Posters}
  \item [01/2024]
    \textbf{American Astronomical Society 243rd Meeting} (\textit{dissertation talk}) \\
    $\cdot$ New Orleans, LA \\
    $\cdot$ \textit{CMB Lensing Measurements with Two Years of Data from the SPT-3G Survey}

  \item [04/2023]
    \textbf{American Physical Society April Meeting 2023} \\
    $\cdot$ Minneapolis, MN \\
    $\cdot$ \textit{CMB Lensing Measurements with Two Years of Data from the SPT-3G Survey}

  \item [05/2022]
    \textbf{Kavli Institute for Particle Physics and Cosmology Tea Talk} \\
    $\cdot$ SLAC National Accelerator Laboratory, Menlo Park, CA \\
    $\cdot$ \textit{Lensing Maps and Transient Science with the South Pole Telescope}

  \item [06/2021]
    \textbf{American Astronomical Society 238th Meeting} (\textit{poster, virtual}) \\
    $\cdot$ \textit{Detection of Stellar Flares at Millimeter Wavelengths with SPT-3G}

  \item [04/2021]
    \textbf{DES Milky Way Working Group Call} (\textit{invited, virtual}) \\
    $\cdot$ \textit{Time-Domain Astronomy with the South Pole Telescope}

  \item [04/2021]
    \textbf{Illinois Astrofest} (\textit{virtual}) \\
    $\cdot$ University of Illinois Urbana-Champaign \\
    $\cdot$ \textit{Time-Domain Astronomy with the South Pole Telescope}

  \item [10/2019]
    \textbf{Center for Astrophysical Surveys Seminar} \\
    $\cdot$ University of Illinois Urbana-Champaign \\
    $\cdot$ \textit{Gravitational Lensing of the CMB: Synergy with Optical Surveys}

  \item[01/2018]
    \textbf{American Astronomical Society 231st Meeting}\\
    $\cdot$ National Harbor, MD\\
    $\cdot$ \textit{Using Vertical Structure to Infer the Total
    Mass Hidden in a Debris Disk}

  \item[7/2017]
    \textbf{Research in Sciences Poster Session}\\
    $\cdot$ Wesleyan University\\
    $\cdot$ Presented poster on AU Mic research.

  \item[10/2016]
    \textbf{Keck Northeast Astronomy Consortium}\\
    $\cdot$ Wesleyan University\\
    $\cdot$ Gave talk on HD 100546 research; published paper in conference proceedings.

  \item[08/2016]
    \textbf{LEAPS Symposium} \\
    $\cdot$ Leiden University \\
    $\cdot$ Gave talk on HD 100546 research to international audience.

  \item[10/2015]
    \textbf{Keck Northeast Astronomy Consortium}\\
    $\cdot$ Williams College\\
    $\cdot$ Gave talk on 49 Ceti research; published paper in conference proceedings.

  \item[07/2015]
    \textbf{Research in Sciences Poster Session}\\
    $\cdot$ Wesleyan University\\
    $\cdot$ Presented poster on 49 Ceti research.

  \end{cvlist}

  \newpage

  \setlength\bibhang{0pt}
  \settowidth{\cvlabelwidth}{1em}%
  \setlength{\cvlabelsep}{0em}

  \begin{cvlist}{Publications}
    \begin{bibunit}
      \nocite{*}
    \item[\textbf{Lead-Author or Substantial Contribution}] \ \\[-1em]
      \putbib[short_author_publications]
    \end{bibunit}
    \begin{bibunit}
      \nocite{*}
    \item[\textbf{Collaboration Papers}] \ \\[-1em]
      \putbib[other_publications]
    \end{bibunit}
  \end{cvlist}

  \cvplace{Gif-sur-Yvette, France}
\end{cv}

\end{document}
\endinput
